Towards a classification framework for social machines

### 2013-02-26 latex/paper_new.tex
pre-final version

Essentially social machines can be characterized as assemblies of manually executed and machine-driven (as in 'automatized') services and the interaction of such services. A  traditional  database,  with  users  looking  up  records  independently  of  each  other  and  an  administrator responsible for the management of the content, has some of the right ingredients, but there is really no social element in this strict provider-consumer relationship. When we fill out a form (e.g., health information, birds spotted in my garden, new construction sites on my way to work) the systems supporting this activity are minimal forms of social machines  because the users  are  part  of  the  'social  computation', in this case the data creation and collection process facilitate through the site.  This social component  becomes  richer when  the database  is  curated by members  of  the broader community (e.g., Wikipedia) and when  the social network adds value implicitly (Amazon) or explicitly (Facebook) to the overall system through their individual or joint activities.

@@

This paper makes two main contributions:

\begin{itemize}
  \item We introduce the notion of 'social machines' and position it in connection to similar ideas and initiatives.
  \item We propose a set of constructs to capture the most important features of social machines and to compare existing instances thereof.
\end{itemize}

The remainder of this paper is organized as follows. In order to pin down what makes up a social machine, and establish the boundaries of this emerging field of research, we need to first understand the relationship between social machines and related topics such as human computation, collective intelligence, and crowdsourcing; Section \ref{sec:comparison} gives a summary of this comparative analysis. In Section \ref{sec:framework} we then present the framework in terms of the methodology used to devise it, the main clusters of constructs used for classification, and an example of using them to compare a selection of popular social machines via grid analysis. We conclude with an outlook of our future development activities, including details about the evaluation of the framework. In addition, we plan to engage with the broader community to set up a 'social machine' bringing together researchers from various disciplines who will contribute to the 'computation' of a joint and widely agreed classification framework for the field.

@@

The Web as a global platform for information access and sharing marked a second essential milestone; in particular, through principles and technologies promoted through Web $2.0$ and the Mobile Web. These developments led to amazing growth in terms of the amounts of content available online and mass participation. They are responsible for hundreds of millions of users all over the globe creating high-quality encyclopedias, publishing Terabytes of multimedia content, contributing to world-class software, and lively taking part in defining the agenda of many aspects of our society. This trend towards 'prosumerism' is finding more and more adopters in the public and private sector as well, as governments and enterprises not only become active in open initiatives, but encourage the participation of their customers and employees in taking decisions related to organizational management, product development, services offers, and policies. In this context, a number of terms are used to refer to the ways people interact with each other and with applications: 'wisdom of the crowds', 'collective intelligence', 'open innovation', 'crowdsourcing', 'human computation', and 'social computing'. They are related, but not synonymous with 'social machines', as explained in the following (see also Figure \ref{socialmachine}).

\begin{figure}[htb]
\begin{center}
\includegraphics[width=0.3\textwidth]{img/socialmachinescope.png}
\caption{Social machines and related areas} \label{fig:social machine}
\end{center}
\end{figure}

@@

\subsection{Methodology}
\label{sec:methodology}
The classification framework presented in this paper was created through knowledge elicitation ~\cite{knowledgeelicitation}. In particular we used the {\it repertory grid} elicitation
technique~\cite{kelly} in order to derive an initial set of {\it elements}, which represent instances of social machines, and {\it constructs}, which capture their most important characteristics.

In repertory grid elicitation a software tool is used to ask users to describe constructs
that differentiate between elements - for example,  prompting the user to create a construct that
differentiates between {\it GalaxyZoo}\footnote{In GalaxyZoo people classify galaxies with collective performance as good as professional astronomers, see \url{http://www.galaxyzoo.org}.} and {\it Facebook}, as two prominent example of social machines our classification framework should apply to. The user describes the opposing poles of the construct - in this case, the user may decide on 'For Science' and 'For Socialising' in order to capture the core distinction between the two system. The user then rates every element with value from $1$ to $5$ on this construct, where $1$ represents an element that is purely
'For Science', and $5$ one that mainly serves  'socializing' purposes. Elements can also be rated with values between the poles if they may correspond to both or neither. The repertory grid software then identifies elements that are the least differentiated when requesting new constructs, in an attempt to elicit as much knowledge as possible from the user. We applied this technique to create an initial set of examples of social machines and to classify them according to specific features. To do so, we asked then computer science researchers familiar with the field to create their own repertory grids, and populate the elements from their own knowledge, and created the constructs using the standard repertory grid elicitation technique described above. This exercise led to $10$ grids, the union of which comprised a total of $56$ unique elements (social machines)
and $117$ different constructs (classifying factors). As the aim of this initial phase was to understand how people perceive the notion of social machine and their most distinctive properties, we allowed the participants to choose the social machines they are familiar with and describe them in their own terms.

@@

\begin{figure}[htb]
\begin{center}
\includegraphics[width=8.5cm]{img/polyarchy.pdf}
\caption{A polyarchy of social machines, illustrating the infrastructure and frameworks used by social machines, and machine-machine usage.} \label{polyarchy}
\end{center}
\end{figure}

@@

The polyarchy is complementary to the classification constructs we elicitated. The constructs apply to all levels of the polyarchy, though some of them might be more easier to understand and describe in the context of some levels. For example, instances of social machines residing at the bottom of the polyarchy are likely to be built with a very concrete audience in mind, and as such the mechanisms to incentivize participation are likely to be clearer and more straightforward to study and adjust than in cases in which such boundaries cannot be defined, such as Facebook, or even Wikipedia. In the same time, these systems will have to decide among an array of diverse basic services and frameworks for their implementations; the impact of such engineering decisions is difficult to predict in great detail as a thorough understanding of what makes certain social machines more successful than others, in terms of technology choices and beyond, is largely missing for most application domains.

@@

\subsection{Constructs}
The clustering process described in Section~\ref{meth} yielded both a
reduction in the overall number of constructs (in particular an
elimination of over-represented areas) and the identification of four
simple themes: sets of constructs pertaining to users and roles within
each machine, second, the nature of tasks, problems and activities
performed through and using the machine, third, motivation for
participation, finally, the contexts of participation.

@@

\begin{table}[htb]
\begin{center}
\begin{tabular}{|p{8cm}|}
\hline

{\bf Tasks, purpose and context of participation} \\
\hline
Activities involve creative production of content \\
Activities involve subjective appraisal of content \\
Activities involve solving (a definable) computation task or set of problems\\
Tasks are domain-specific \\
The machine owner (service/platform) derives benefit from participation \\
Activities and tasks are pre-defined or participant-defined \\
Variation in types of contributions and tasks \\
Participants' participate to fulfill needs of a role  \\
\hspace{1cm} work-related/professional role \\
\hspace{1cm} home/family related \\
\hspace{1cm} social role \\
\hspace{1cm} leisure/entertainment role \\
Participation is done via: \\
\hspace{1cm} mobile devices \\
\hspace{1cm} web browsers \\
\hspace{1cm} apps \\
\hspace{1cm} sensors/sensing (location sensing and wearable devices) \\
Participation is done in a mobile context \\
\hspace{1cm} Physical location is relevant to the service \\
\hline
{\bf Participants, roles and incentives} \\
\hline
Generality of audience \\
Participant autonomy \\
Participant anonymity \\
Extent of hierarchical organization of roles \\
Clear separation of roles/responsibilities among participants \\
Participants are intrinsically motivated:\\
\hspace{1cm} to gain/share knowledge \\
\hspace{1cm} to ``get something done'' \\
\hspace{1cm} to ``for fun/entertainment''\\
\hspace{1cm} to ``be social'' \\
\hspace{1cm} for the benefit of a specific group of people who need help \\
\hspace{1cm} for the benefit of society as a whole \\
Participation is motivated by extrinsic reward (payment, status etc) \\
\hline

\end{tabular}
\end{center}
\caption{Consolidated constructs of social machines.} \label{table:constructs}
\end{table}

\subsubsection{System view: tasks, purpose and context of participation}
The first set of constructs pertain to \emph{what} the participants
do, both individually within the system, and collectively as a whole.
Individual activities might include creation of content, subjective
appraisal of existing content, posing of problems, solving of
problems, andf so on.  The resulting computation, looked at from a
macroscopic whole, might be the identification of high-quality
artistic content, insight, or creative works, collective problem
solving, and so on.  Such tasks might be pre-defined by the system
designers, or brought to the system by participants.  Futher, these
tasks might be of a specific type or class, or of many varied types or
encompass a general class of activities.  To capture these
difeferences, we include a construct that differentiates machines with
pre-defined problem spaces or are open to participant-provided tasks
and activities.  Finally, we ask whether the benefit derived by
participants is distinct from that gained by machine/platform owners
and service providers.

The final set of constructs pertains to the context(s) of each
participant's greater (life-) activities that participation is
performed.  This context could be in a work context, home/family
context, for leisure, and at a desk or away from a computer (in a
mobile context).  PArticipation could be conducted via apps, web
browsers, sensors and other technolgoies.  Due to the predominance of
``geo-locative'' services that incorporate information about the
user's location, we have included a construct identifying the degree
to which the user's location is used by the machine.

\subsubsection{Social view: users, roles, motivation and incentives}

The first set of constructs pertain to the human participants of
social machines and the way(s) participants are organised within them.
One construct identifies whether participants constitute members of
the general public, or are of a specific demongraphic, external group
or organisation, occupation, or of a particular expertise.  Two
constructs, \emph{autonomy} and \emph{anonymity} pertain to the degree
to which interaction is constrained among participants, and to the
degree to which participants' identities are used and disclosed within
the machine. The final two constructs pertain to the roles: first,
as to whether multiple roles are pre-differentiated by the machine or
whether all participants essentially initially assume the same role,
second whether roles (pre-defined or emergent) are archical.

The third set of themes pertains to the motivation that sustains
continued participation in these systems.  We identiy 6 different
kinds of common sources of intrinsic motivation; first, that
participation is fun, second that it accomplishes an activity that the
participant enjoys or wanted to get done, third, that it satisfies the
desire to gain or share knowledge, or, fourth, or to be social.  The
last two pertain to philanthropy, whether participation is seen to
benefit a particular group of disadvantaged people or individuals who
need assistance, or, finally, if participation is beneficial to
society as a whole.  Finally, we ask whether extrinsic motivational
factors also contributed to sustained participation - such factors
could include money reward, status, recognition and so on

@@

\subsection{Using the classification framework}
\label{sec:usage}
In order to demonstrate usage of our framework's constructs, we ranked our social machines
by their Alexa ranking, and performed a full repertory grid elicitation exercise over them.
This process involved ranking each social machine against the constructs, rating each construct
from 1 to 5. In total, $680$ ranking were made, and the
results are illustrated in a grid and accompanying dendrograms for element and construct
similarity in Figure~\ref{dendrogram}.

\begin{figure}[htb]
\begin{center}
\includegraphics[width=4cm]{img/dendrogram-elements.png}
\caption{Dendrogram of elements (social machines) from a repertory grid exercise of the top 20 (ranked by Alexa) social machines from our element set, against our consolidated constructs.} \label{dendrogram}
\end{center}
\end{figure}

\begin{figure}[htb]
\begin{center}
\includegraphics[width=8cm]{img/dendrogram-constructs.png}
\caption{Dendrogram of constructs from a repertory grid exercise of the top 20 (ranked by Alexa) social machines from our element set, against our consolidated constructs.} \label{dendrogram}
\end{center}
\end{figure}

From the dendrograms we can get a sense of similarities and separation of social machines. For
instance, it is clear that YouTube, Vimeo, Reddit and Digg are similar, which is logical, because
they are all social machines that focus on sharing content. Likewise, Quora and StackOverflow
are also similar, and again, this makes sense, because they are both question answering social
machines. There is also a dendrogram for the constructs, to show correlations between them. This
dendrogram indicates that there is a correlation between the ``Single Participant Roles'' and
``No Hierarchy in Participant Roles'' constructs, which is because they refer to the same
aspect of social machines. Another, but less obvious similarity is between the ``Large variety of
social features'' and ``Benefit NOT for the world/society'', which indicates that of the social
machines we have rated, there is a correlation between social features and social machines that
do not benefit the world at large, and vice-versa.

We envisage two key initial uses for this framework: for classifying social machines; and for
predicting behaviour of new/future social machines. For classifying social machines, it will
enable direct comparisons to be made with existing machines, as well as provide a platform
for discussions about social machines. We also think that the data from a large-scale classification
exercise may be useful for researchers who wish to apply prediction techniques (such as SVMs)
to social machines, their behaviour and impact.

\section{Future work: building a social machine around the classification framework}

\section{Acknowledgments}

This work is supported under SOCIAM: The Theory and Practice of Social Machines.  The SOCIAM Project is funded by the UK Engineering and Physical Sciences Research Council (EPSRC) under grant number EP/J017728/1 and comprises the Universities of Southampton, Oxford and Edinburgh.

>>>>

Essentially social machines can be characterized as assemblies of manually executed and machine-driven (as in 'automatized') services and the interaction of such services. A  traditional  database,  with  users  looking  up  records  independently  of  each  other  and  an  administrator responsible for the management of the content, has some of the right ingredients, but there is really no social element in this strict provider-consumer relationship. When we fill out a form (e.g., health information, birds spotted in my garden, new construction sites on my way to work) the systems supporting this activity are minimal forms of social machines  because the users  are  part  of  the  'social  computation', in this case the data creation and collection process facilitate through the site.  This social component  becomes  richer when  the database  is  curated by members  of  the broader community (e.g., Wikipedia) and when  the social network adds value implicitly (e.g., Amazon) or explicitly (e.g., Facebook) to the overall system through their individual or joint activities.

@@

This paper makes two main contributions: First, we introduce the notion of 'social machines' and position it in connection to similar ideas and initiatives. Second we propose a set of constructs to capture the most important features of social machines and to compare existing instances thereof.

The remaining sections are organized as follows. In order to pin down what makes up a social machine, and establish the boundaries of this emerging field of research, we need to first understand the relationship between social machines and related topics such as human computation, collective intelligence, and crowdsourcing; Section \ref{sec:comparison} gives a summary of this comparative analysis. In Section \ref{sec:framework} we then present the framework in terms of the methodology used to devise it, the main clusters of constructs used for classification, and an example of using them to compare a selection of popular social machines via grid analysis. We conclude with an outlook of our future development activities, including details about the evaluation of the framework. In addition, we plan to engage with the broader community to set up a 'social machine' bringing together researchers from various disciplines who will contribute to the 'computation' of a joint and widely agreed classification framework for the field.

@@

The Web as a global platform for information access and sharing marked a second essential milestone; in particular, through principles and technologies promoted through Web $2.0$ and the Mobile Web. These developments led to amazing growth in terms of the amounts of content available online and mass participation. They are responsible for hundreds of millions of users all over the globe creating high-quality encyclopedias, publishing Terabytes of multimedia content, contributing to world-class software, and lively taking part in defining the agenda of many aspects of our society. This trend towards 'prosumerism' is finding more and more adopters in the public and private sector as well, as governments and enterprises not only become active in open initiatives, but encourage the participation of their customers and employees in taking decisions related to organizational management, product development, services offers, and policies. In this context, a number of terms are used to refer to the ways people interact with each other and with applications: 'wisdom of the crowds', 'collective intelligence', 'open innovation', 'crowdsourcing', 'human computation', and 'social computing'. They are related, but not synonymous with 'social machines', as explained in the following (see also Figure \ref{fig:socialmachine}).

\begin{figure}[htb]
\begin{center}
\includegraphics[width=0.25\textwidth]{img/socialmachinescope.png}
\caption{Social machines and related areas} \label{fig:social machine}
\end{center}
\end{figure}

@@

\subsection{Methodology}
\label{sec:methodology}
The classification framework presented in this paper was created through knowledge elicitation ~\cite{knowledgeelicitation}. In particular we used the {\it repertory grid} elicitation
technique~\cite{kelly} in order to derive an initial set of {\it elements}, which represent instances of social machines, and {\it constructs}, which capture their most important characteristics.
In repertory grid elicitation a software tool is used to ask users to describe constructs
that differentiate between elements - for example,  prompting the user to create a construct that
differentiates between {\it GalaxyZoo}\footnote{In GalaxyZoo people classify galaxies with collective performance as good as professional astronomers, see \url{http://www.galaxyzoo.org}.} and {\it Facebook}, as two prominent example of social machines our classification framework should apply to. The user describes the opposing poles of the construct - in this case, the user may decide on 'For Science' and 'For socialising' in order to capture the core distinction between the two system. The user then rates every element with value from $1$ to $5$ on this construct, where $1$ represents an element that is purely
'For Science', and $5$ one that mainly serves  'socializing' purposes. Elements can also be rated with values between the poles if they may correspond to both or neither. The repertory grid software then identifies elements that are the least differentiated when requesting new constructs, in an attempt to elicit as much knowledge as possible from the user. We applied this technique to create an initial set of examples of social machines and to classify them according to specific features. To do so, we asked then computer science researchers familiar with the field to create their own repertory grids, and populate the elements from their own knowledge, and created the constructs using the standard repertory grid elicitation technique described above. This exercise led to $10$ grids, the union of which comprised a total of $56$ unique elements (social machines)
and $117$ different constructs (classifying factors). As the aim of this initial phase was to understand how people perceive the notion of social machine and their most distinctive properties, we allowed the participants to choose the social machines they are familiar with and describe them in their own terms.

@@

\begin{figure}[htb]
\begin{center}
\includegraphics[width=0.4\textwidth]{img/polyarchy.pdf}
\caption{A polyarchy of social machines, illustrating the infrastructure and frameworks used by social machines, and machine-machine usage} \label{polyarchy}
\end{center}
\end{figure}

@@

The polyarchy is complementary to the classification constructs we elicited. The constructs apply to all levels of the polyarchy, though some of them might be more easier to understand and describe in the context of some levels. For example, instances of social machines residing at the bottom of the polyarchy are likely to be built with a very concrete audience in mind, and as such the mechanisms to incentivize participation are likely to be clearer and more straightforward to study and adjust than in cases in which such boundaries cannot be defined, such as Facebook, or even Wikipedia. In the same time, these systems will have to decide among an array of diverse basic services and frameworks for their implementations; the impact of such engineering decisions is difficult to predict in great detail as a thorough understanding of what makes certain social machines more successful than others, in terms of technology choices and beyond, is largely missing for most application domains.

@@

\subsection{Constructs}
The clustering process described in Section~\ref{meth} yielded both a
reduction in the overall number of constructs (in particular an
elimination of over-represented areas) and the identification of three
simple themes: a

sets of constructs pertaining to users and roles within
each machine, second, the nature of tasks, problems and activities
performed through and using the machine, third, motivation for
participation, finally, the contexts of participation.

@@

\begin{table}[htb]
\begin{center}
\begin{scriptsize}
\begin{tabular}{|p{8cm}|}
\hline
{\bf Tasks, purpose and context of participation} \\
\hline
Activities involve creative production of content \\
Activities involve subjective appraisal of content \\
Activities involve solving (a definable) computation task or set of problems\\
Tasks are domain-specific \\
The machine owner derives benefit from participation \\
Activities and tasks are pre-defined or participant-defined \\
Variation in types of contributions and tasks \\
Participants' participate to fulfill needs of a role  \\
\hspace{1cm} work-related/professional role \\
\hspace{1cm} home/family related \\
\hspace{1cm} social role \\
\hspace{1cm} leisure/entertainment role \\
Participation is done via: \\
\hspace{1cm} mobile devices \\
\hspace{1cm} Web browsers \\
\hspace{1cm} apps \\
\hspace{1cm} sensors/sensing (location sensing and wearable devices) \\
Participation is done in a mobile context \\
\hspace{1cm} Physical location is relevant to the service \\
\hline
{\bf Participants and roles} \\
\hline
Generality of audience \\
Participant autonomy \\
Participant anonymity \\
Extent of hierarchical organization of roles \\
Clear separation of roles among participants \\
{\bf Motivation and incentives} \\
\hline
Participants are intrinsically motivated:\\
\hspace{1cm} to gain/share knowledge \\
\hspace{1cm} to ``get something done'' \\
\hspace{1cm} to ``for fun/entertainment''\\
\hspace{1cm} to ``be social'' \\
\hspace{1cm} for the benefit of a specific group of people who need help \\
\hspace{1cm} for the benefit of society as a whole \\
Participation is motivated by extrinsic reward (payment, status) \\
\hline

\end{tabular}
\end{scriptsize}
\end{center}
\caption{Consolidated constructs of social machines} \label{table:constructs}
\end{table}

\subsubsection{Tasks, purpose, and context of participation}
The first set of constructs pertain to \emph{what} the participants
do, both individually within the system, and collectively as a whole.
Individual activities might include creation of content, subjective
appraisal of existing content, posing of problems, solving of
problems, and so on.  The resulting 'computation', looked at from a
macroscopic whole, might be the identification of high-quality
artistic content, insight, or creative works, collective problem
solving, and so on.  Such tasks might be pre-defined by the system
designers, or brought to the system by participants.  Further, these
tasks might be of a specific type or class, or of many varied types or
encompass a general class of activities.  To capture this diversity, 
we include a construct that differentiates machines with pre-defined problem spaces or are open to participant-provided tasks and activities.  Finally, we ask whether the benefit derived by
participants is distinct from that gained by machine/platform owners
and service providers.

The second set of constructs pertains to the context(s) of each participant's greater (life-) activities in which the interaction with the system occurs.  This context could be in a work context, home/family context, for leisure, and at a desk or away from a computer (in a mobile context).  Participation could be conducted via apps, Web browsers, sensors and other technologies.  Due to the predominance of geolocation services, one of the constructs collected through knowledge elicitation identifies the degree to which the user's location is used by the machine.

\subsubsection{Participants and roles}
This set of constructs pertain to the human participants of social machines and the way(s) participants are organized within them. One construct identifies whether participants constitute members of the general public, or are of a specific demographic, external group or organization, occupation, or of a particular expertise.  Two constructs, \emph{autonomy} and \emph{anonymity} pertain to the degree to which interaction is constrained among participants, and to the degree to which participants' identities are used and disclosed within the machine. Additional two constructs pertain to the roles: first, as to whether multiple roles are pre-differentiated by the machine or
whether all participants essentially initially assume the same role, second whether roles (pre-defined or emergent) are hierarchical.

\subsubsection{Motivation and incentives}
Related to the social structure  motivation that sustains
continued participation in these systems.  We identify six different
kinds of common sources of intrinsic motivation; first, that
participation is fun, second that it accomplishes an activity that the
participant enjoys or wanted to get done, third, that it satisfies the
desire to gain or share knowledge, or, fourth, or to be social.  The
last two pertain to philanthropy, whether participation is seen to
benefit a particular group of disadvantaged people or individuals who
need assistance, or, finally, if participation is beneficial to
society as a whole.  Finally, we ask whether extrinsic motivational
factors also contributed to sustained participation - such factors
could include money reward, status, recognition and so on \cite{kuznetsov2006motivations,lakhani2003hackers,tagging}.
%Peer-to-peer filesharing~\cite{p2p}; Tagging~\cite{tagging}; and Virtual
%Communities~\cite{ardichvili2003motivation,Moore:2007:UMM:1235000.1235035}.

@@

\subsection{Using the classification framework}
\label{sec:usage}
In order to demonstrate usage of the constructs we ordered our social machines
by their Alexa ranking, and performed a full repertory grid elicitation exercise over the first $20$ of the $56$ collected in the first phase of the methodology (see Section \ref{sec:methodology}). This process assessing each social machine against the constructs by choosing a rate between $1$ and $5$. This led to a total of $680$ ratings, and the
results are illustrated in a grid and accompanying dendrograms for element and construct
similarity in Figure~\ref{dendrogram}.

\begin{figure}[htb]
\begin{center}
\includegraphics[width=0.2\textwidth]{img/dendrogram-elements.png}
\caption{Dendrogram of elements (social machines) from a repertory grid exercise of the top $20$ (ranked by Alexa) social machines from our element set, against our consolidated constructs} \label{dendrogram}
\end{center}
\end{figure}

\begin{figure}[htb]
\begin{center}
\includegraphics[width=0.4\textwidth]{img/dendrogram-constructs.png}
\caption{Dendrogram of constructs from a repertory grid exercise of the top $20$ (ranked by Alexa) social machines from our element set, against our consolidated constructs} \label{dendrogram}
\end{center}
\end{figure}

From the dendrograms we can get a sense of similarities and separation of social machines. For
instance, it is clear that YouTube, Vimeo, Reddit, and Digg are similar, which is not unexpected, due to their focus on sharing content. Likewise, the exercise confirms commonalities between Quora and StackOverflow, which can be easily understood, given their question answering purpose. There is also a dendrogram for the constructs, which can be used, as more data is collected, as an indicator for correlation between constructs. In this particular case, dendrogram suggests that there is a correlation between the 'Single Participant Roles' and 'No Hierarchy in Participant Roles' constructs, which is because they refer to the same aspect of social machines. Another, but less obvious connection is between the 'Large variety of social features' and 'Benefit NOT for the world/society', which requires further investigation.

We envisage three key use cases for our framework. First and foremost the framework forms the basis for a coherent and comprehensive description and classification of social machines. The knowledge elicitation technique referred to in this paper is just one of the many tools which can assist this exercise. In addition, it enables direct comparisons between systems, and provides a platform for terminologically consistent discussions about the field as a whole and specific instantiations. Finally, we also think that with more and more classification data becoming available, for instance, also through functionality that automatically monitors and computes the values of the some of the constructs introduced earlier, researchers will wish to apply prediction techniques (such as SVMs) to social machines, their behavior, and impact.

\section{Future work: building a social machine around the classification framework}
In this paper we presented a first outline of a possible organization structure for the design space of social machines. We introduced the key features of the notion of social machines, and explained how this notion differs from related areas such as collective intelligence, crowdsourcing and human computing. We then discussed the basic constructs of a classification framework created through the repertory grid elicitation technique. In its current version the framework consists of $31$ constructs clustered according to the main components of social machines: social and machine-driven services, and the interactions between them. Our future work will mostly be dedicated to the revision of this initial list of constructs. Most notably, we would like to add a number of constructs to emphasize the importance of the interaction aspect in the theory, engineering, and operation of social machines. This includes constructs related to the human-computer workflows supported (e.g., human intelligence replacing automatic algorithms as in GWAPs vs the different bots available on Wikipedia complementing collaborative manual editing), the way inputs from the crowd are validated, aggregated and turned into a useful outcome, as well as issues related collaboration and coordination, and interface and communication design. In addition, we plan to engage with the broader researcher community by setting up a social machine to facilitate discussions and feedback collection. One specific idea that is currently under development is to build a microtask-like environment, including specific game elements, in which participants are asked to provide answers to atomic challenges that rate and compare a pair of social machine instances according to a construct of our framework.

\section{Acknowledgments}
This work is supported under SOCIAM: The Theory and Practice of Social Machines.  The SOCIAM Project is funded by the UK Engineering and Physical Sciences Research Council (EPSRC) under grant number EP/J017728/1 and comprises the Universities of Southampton, Oxford and Edinburgh.
